Probability

### 2016-04-24 prob.tex
prob.tex \section{Quantum Probability Spaces} update...

Consider the two-qubit Hilbert space with computational basis
$\ket{0}$ and $\ket{1}$ and states:
\[
\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}, \qquad\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}\,.
\]
The set of events associated with this Hilbert space consists of all
projections including the empty projection $\mathbb{0}$ and the unit
projection $\mathbb{1} = \proj{0}+\proj{1}$:
\[ 
\{ \mathbb{0}, \proj{0}, \proj{1}, \proj{\ps}, \proj{\ms}, \ldots, \mathbb{1}\}
\]
Each event is interpreted as a possible post post-measurement state of
a quantum system as follows: given some arbitrary current quantum state
$\ket{\psi}$ to be measured, the event $\proj{0}$ states that the
post-measurement state will be $\ket{0}$; the event $\proj{1}$ states
that the post-measurement state will be $\ket{1}$; the event
$\proj{\ps}$ states that the post-measurement state will be
$\ket{\ps}$; the event $\proj{\ms}$ states that the post-measurement
state will be $\ket{\ms}$; the event $\mathbb{1}$ states that the
post-measurement state will be a linear combination of $\ket{0}$ and
$\ket{1}$; and the event $\mathbb{0}$ states that the post-measurement
state will be the empty state.

Irrespective of the current state $\ket{\psi}$ and
irrespective of the particular experiment, the probability of event
$\mathbb{0}$ will always be 0 (it is an impossible event) and the
probability of event $\mathbb{1}$ will always be 1 (it is a certain
event). The probabilities attached to other events will depend on the
particular state in question. If the state is $\ket{0}$, the
probability of event $\proj{0}$ is 1; the probability of event
$\proj{1}$ is 0; the probability of event $\proj{\ps}$ is
$\frac{1}{2}$; and the probability of event $\proj{\ms}$ is
$\frac{1}{2}$. If the state is $\ket{\ps}$, the probability of each
event $\proj{0}$ and $\proj{1}$ will be $\frac{1}{2}$; the
probability of event $\proj{\ps}$ is 1; and the probability of event
$\proj{\ms}$ is 0.

@@

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)= \ip{\psi}{E\psi} \label{eq:Born}
\end{equation}
which is called the Born rule~\cite{Born1984,Mermin2007}.\\
\yutsung{One small thing is whether we should say
\begin{quote}
for each normalized state~$\ket{\psi}$...$\mu_{\psi}(E)=\ip{\psi}{E\psi}$
\end{quote}
as in Mermin~\cite{Mermin2007} and Maassen~\cite{Maassen2010}
or
\begin{quote}
for each state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $...$\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}$
\end{quote}
as in Abramsky~\cite{DBLP:journals/corr/abs-0910-2393}. Personally,
I prefer Abramsky's approach, because normalizing a state in DQT is
messier than in CQT, and we have already considered states whose norm
is not one in our previous paper~\cite{DQT2014}.}\\
Our previous examples can be verified in terms of $\mu_{\psi}$. Because
$\proj{0}$ and $\proj{-}$ are non-commutative, their total probabilities
are not a probability of an event
\begin{eqnarray*}
\mu_{0}(\proj{0})+\mu_{0}(\proj{-}) & = & \frac{3}{2}>1\textrm{ .}
\end{eqnarray*}
When events are orthogonal, i.e., commutative, they can be measured
together and their probability is additive: 
\[
\mu_{0}(\proj{0})+\mu_{0}(\proj{1})=\mu_{0}(\proj{+})+\mu_{0}(\proj{-})=1=\mu_{0}(\mathbb{1})\textrm{ .}
\]
Similarly to the classical case, the total probability is one and
the orthogonal additivity can be summarized into the following two
rules:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_{i}$, we have $\mu(\sum_{i}E_{i})=\sum_{i}\mu(E_{i})$.
\end{itemize}
These two conditions define quantum probability measures. Recall that
the Born rule (\ref{eq:Born}) induces a probability measure from
a state. Naturally, we may ask: is every probability measure induced
from a state by the Born rule? The answer is yes by Gleason's theorem
when the dimension~$\ge3$ and extend the domain of the Born rule
to mixed state~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.

>>>>

Consider the one-qubit Hilbert space with computational basis
$\ket{0}$ and $\ket{1}$ and states:
\[
\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}, \qquad\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}\,.
\]
The set of events associated with this Hilbert space consists of all
projections including the empty projection $\mathbb{0}$ and the unit
projection $\mathbb{1} = \proj{0}+\proj{1}$:
\[ 
\{ \mathbb{0}, \proj{0}, \proj{1}, \proj{\ps}, \proj{\ms}, \ldots, \mathbb{1}\}
\]
Each event is interpreted as a possible post-measurement state of
a quantum system as follows: given some arbitrary current quantum state
$\ket{\psi}$ to be measured, the event $\proj{0}$ states that the
post-measurement state will be $\ket{0}$; the event $\proj{1}$ states
that the post-measurement state will be $\ket{1}$; the event
$\proj{\ps}$ states that the post-measurement state will be
$\ket{\ps}$; the event $\proj{\ms}$ states that the post-measurement
state will be $\ket{\ms}$; the event $\mathbb{1}$ states that the
post-measurement state will be a linear combination of $\ket{0}$ and
$\ket{1}$; and the event $\mathbb{0}$ states that the post-measurement
state will be the empty state.

In classical probability space, two particular coins can define a
probability measure on every event. Similarly, a quantum system can
define a quantum probability measure on its event. The difference
is that only orthogonal events can test together. For example, asking
whether event~$\mathbb{1}$ occurs is the same as asking whether
event~$\proj{0}$ or event~$\proj{1}$ occurs simultaniously. However,
we can never ask whether event~$\proj{0}$ or event~$\proj{-}$
occurs simultaniously. Similarly to the classical case, a \emph{quantum
probability space}~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}
can be summarized into the following two rules\footnote{Recently, people extend the domain of $\mu$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_{i}$, we have $\mu\left(\sum_{i}E_{i}\right)=\sum_{i}\mu\left(E_{i}\right)$. 
\end{itemize}
\begin{example}{[}One-qubit Hilbert space{]} Consider an experiment
that with a qubit. The events are discribed previously. Depending on
the systems, we can define different quantum probability measures.
Here is a possible one: 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\mu_{0}\left(\mathbb{0}\right) & = & 0\\
\mu_{0}\left(\proj{0}\right) & = & 1\\
\mu_{0}\left(\proj{1}\right) & = & 0\\
 & \vdots
\end{array} & \begin{array}{rcl}
\mu_{0}\left(\mathbb{1}\right) & = & 1\\
\mu_{0}\left(\proj{\ps}\right) & = & 1/2\\
\mu_{0}\left(\proj{\ms}\right) & = & 1/2\\
 & \vdots
\end{array}\end{array}
\]
The assignment satisfies the two constraints for quantum probability
measures: the probability of $\mathbb{1}$ is 1, and the probability
of every summation of orthogonal events (e.g., $\proj{0}+\proj{1}$)
is the sum of the individual probabilities. \end{example}

@@

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)= \ip{\psi}{E\psi} \label{eq:Born}
\end{equation}
which is called the Born rule~\cite{Born1984,Mermin2007}.\\
\yutsung{One small thing is whether we should say
\begin{quote}
for each normalized state~$\ket{\psi}$...$\mu_{\psi}(E)=\ip{\psi}{E\psi}$
\end{quote}
as in Mermin~\cite{Mermin2007} and Maassen~\cite{Maassen2010}
or
\begin{quote}
for each state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $...$\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}$
\end{quote}
as in Abramsky~\cite{DBLP:journals/corr/abs-0910-2393}. Personally,
I prefer Abramsky's approach, because normalizing a state in DQT is
messier than in CQT, and we have already considered states whose norm
is not one in our previous paper~\cite{DQT2014}.}\\
These two conditions define quantum probability measures. Recall that
the Born rule (\ref{eq:Born}) induces a probability measure from
a state. Naturally, we may ask: is every probability measure induced
from a state by the Born rule? The answer is yes by Gleason's theorem
when the dimension~$\ge3$ and extend the domain of the Born rule
to mixed state~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}.
